Ontology Matching with word2vec

### 2016-02-26 Paper/Draft.tex
Added Combining edges with nodes, hot start

 \documentclass{article}
 \usepackage[utf8]{inputenc}
 \usepackage[titletoc]{appendix}
 \title{Ontology Matching with word2vec}
 \author{Harmen Prins}

@@

 \subsection{Theory}
 \subsection{Possibilities} % Rename please
  \subsubsection{Training corpus}
  Word embedding models need to be trained on a large corpus. These corpora need to cover the concepts that are contained in the ontology, but also need to be large enough to build good embeddings. Since the ontologies themselves do not necessarily contain enough examples to embed the concepts properly, I need to use ontologies that are likely to contain the concepts I want to align.
  As a benchmark, I will use the WordNet and NELL ontologies. To check if adding the ontologies that are to be aligned helps, I will also train a model on an ontology consisting of WordNet, NELL and the two alignment candidates.
  \subsubsection{Node vectorisation}
  The most basic node2vec model converts a node to a vector purely based on its own label. This method has one advantage over simple string matching: if no node is found in the other ontology that matches exactly, we can still find a node that is similar since it is close in the vector space. Therefore, this method should already be an improvement over the most basic string matching algorithms. It may even be competitive with more advanced string matchers (that look at substrings of labels). For example, \emph{ear lobe} and \emph{ear} will be matched by a substring matcher, but are also semantically similar since they will often be mentioned in the same context.
  \subsubsection{Combine with neighbours}
  To improve the model, we can add context information from the neighbours of a node. The most basic context adaptation would be to create a vector as the average of the context vectors. The resulting vector should be combined with the node vector of the node that is being investigated. This can be done by averaging or weighted averaging where the context is weighed more if there are more neighbours, though not necessarily linearly. These possibilities were also investigated.
  \subsubsection{Select from senses}
  Another way to take into account the influence of neighbouring nodes is to train a multi-sense node embedding model as described in {appropriate section}, and selecting from the different node senses based on which is closest to the context average.
  \subsubsection{Adding edge labels}
  To add more information to the model, edge labels can also be added to the context of a node. This method effectively doubles the training data and context size when creating the context average. This should make the model more robust, although edge labels may be duplicate (one has many unique family members) and less informative in general, so a lower weight may be appropriate. Another problem with this method is that the relation between a neighbour node and its corresponding edge is lost, since they are just treated as independent contexts.
  \subsubsection{Select from senses}
  This method is exactly the same as the selecting from node senses with just the neighbour nodes except that you now also take into account the edge labels.
  \subsubsection{Combining edges with nodes}
  As can be read in {appropriate section}, 
  \subsubsection{Hot start}
 \subsection{Discarded possibilities} % Rename
  \subsubsection{Training on text}
  \subsubsection{Just words}
  \subsubsection{Just edges}
  \subsubsection{Context matrix representation}
 \subsection{Implementation}

>>>>

 \documentclass{article}
 \usepackage[utf8]{inputenc}
 \usepackage[titletoc]{appendix}
 \usepackage{comments}
 \title{Ontology Matching with word2vec}
 \author{Harmen Prins}

@@

 \subsection{Theory}
 \subsection{Possibilities} % Rename please
  \subsubsection{Training corpus}
  Word embedding models need to be trained on a large corpus. These corpora need to cover the concepts that are contained in the ontology, but also need to be large enough to build good embeddings. Since the ontologies themselves do not necessarily contain enough examples to embed the concepts properly, I need to use ontologies that are likely to contain the concepts I want to align.
  As a benchmark, I will use the WordNet and NELL ontologies. To check if adding the ontologies that are to be aligned helps, I will also train a model on an ontology consisting of WordNet, NELL and the two alignment candidates.
  \subsubsection{Node vectorisation}
  The most basic node2vec model converts a node to a vector purely based on its own label. This method has one advantage over simple string matching: if no node is found in the other ontology that matches exactly, we can still find a node that is similar since it is close in the vector space. Therefore, this method should already be an improvement over the most basic string matching algorithms. It may even be competitive with more advanced string matchers (that look at substrings of labels). For example, \emph{ear lobe} and \emph{ear} will be matched by a substring matcher, but are also semantically similar since they will often be mentioned in the same context.
  \subsubsection{Combine with neighbours}
  To improve the model, we can add context information from the neighbours of a node. The most basic context adaptation would be to create a vector as the average of the context vectors. The resulting vector should be combined with the node vector of the node that is being investigated. This can be done by averaging or weighted averaging where the context is weighed more if there are more neighbours, though not necessarily linearly. These possibilities were also investigated.
  \subsubsection{Select from senses}
  Another way to take into account the influence of neighbouring nodes is to train a multi-sense node embedding model as described in {appropriate section}, and selecting from the different node senses based on which is closest to the context average.
  \subsubsection{Adding edge labels}
  To add more information to the model, edge labels can also be added to the context of a node. This method effectively doubles the training data and context size when creating the context average. This should make the model more robust, although edge labels may be duplicate (one has many unique family members) and less informative in general, so a lower weight may be appropriate. Another problem with this method is that the relation between a neighbour node and its corresponding edge is lost, since they are just treated as independent contexts.
  \subsubsection{Select from senses}
  This method is exactly the same as the selecting from node senses with just the neighbour nodes except that you now also take into account the edge labels.
  \subsubsection{Combining edges with nodes}
  As can be read in {appropriate section}, it is possible to drop the bag-of-words assumption that the context is sequence invariant. \begin{comment}maybe ref too\end{comment} This means we can for example have the context be (previous word, next word) and those words will be treated differently. Similarly we can separate the edge and node labels and treat them differently. This ensures that the model will find any relationship between the edge and the node if it exists and will take this into account. For example, the system might use just the information 'isMarriedTo' to infer that someone is a human, but may infer that 'isMarriedTo' in combination with a man usually refers to a woman. \begin{comment} Ja doe maar een ander voorbeeld ivm gelijkheid e.d.\end{comment}
  \subsubsection{Hot start}
  It is hard to classify the algorithm in terms of the normal ontology alignment methods. It is not purely string matching, since it takes into account information from the context. However, it is also not structural, as it can work without a seed alignment and uses information other than the structure as well. It is not logical \begin{comment} Goeie woord?\end{comment} since it does not use inference. Nor is it terminological, which looks at dictionaries or other ontologies to find matches. However, it does use other ontologies (or text corpi\begin{comment}Check spelling\end{comment}) implicitly, so if it must be grouped it would be terminological. \begin{comment}Moet voorafgaand stukje op een andere plek?\end{comment}
  The algorithm can find matches based purely on the given node and edge labels. However the algorithm might benefit from a hot start. Such a seed alignment might allow the algorithm to train with certain words that are known synonyms as if they were the same word, thus increasing the number of training samples per word (on known relevant words) and decreasing sparsity. This may help the model, improving performance. However, since some information from other algorithms is now used, it would not be fair to compare the results to the cold start algorithm results. For example, it may just add the results from the generic string matcher to its own results, improving performance, without learning any new relationships. A new testing method had to be designed for this algorithm. This method would have to compare it to the string matchers, to see if it improved over their results, rather than compare it to the cold start algorithm. However, if we compare the cold and hot algorithms with the generic string matchers, we may be able to compare them indirectly. \begin{comment}Dit moet weg als het bs blijkt te zijn\end{comment}
 \subsection{Discarded possibilities} % Rename
  \subsubsection{Training on text}
  Rather than training on ontologies, it would be possible to train the model on text. Since there is more text available than ontologies when counting the number of training samples, it would make sense to use a purely text-trained model. However, this method has downsides which made me decide not to use it. Namely, the context of a word in a text corpus is very different from a node in an ontology. A word is surrounded by the words directly around it, or the sentence it is in, or the paragraph or entire document. However all these contexts are just words, usually represented as a bag of words to ensure enough data for the model. In an ontology, the context is a combination of a node with an edge, which have a very different relation than merely co-occurring. The relation may be inferred from one sentence, or many different documents. \begin{comment}Herschrijf dit aub want het is niet goed\end{comment}
  \subsubsection{Just words}
  \subsubsection{Just edges}
  \subsubsection{Context matrix representation}
 \subsection{Implementation}
